Automatic conversion of colloquial Finnish to standard Finnish

### 2015-01-26 finnormal-nodalida-2015/finnormal2015.tex
more stuff
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58610 72bbbca6-d526-0410-a7d9-f06f51895060

% supervised systems trained on the normalised corpus outperform those trained on the standardised corpus (WHY?)

\begin{table}
  \centering
  \begin{tabular}{|l|r|r|r|}
     \hline
    \textbf{System} & \textbf{PER} & \textbf{WER} & \textbf{BLEU} \\
     \hline
     Colloquial   & 45.31 & 71.28 & 11.95 \\
     Normalised   & 34.63 & 62.34 & 21.81 \\
     Standardised & 38.67 & 63.35 & 17.62 \\
     \hline
     rules-2      & 42.71 & 70.42 & 12.28 \\
     \hline
     cbmt-cn      & 44.44 & 74.17 & 12.35 \\
     pbmt-cn      & \textbf{38.38} & \textbf{66.81} & \textbf{17.90} \\
     \hline
  \end{tabular}
  \caption{Results for the Finnish to English translation task. The first three rows are the results 
    from translating the sections of the corpus. As with the }
  \label{table:results-trad}
\end{table}

% example translations

@@

We have shown that converting from colloquial Finnish to standard Finnish substantially
helps with the machine translation task.

%The main contribution of this paper is a parallel corpus of colloquial Finnish
%and standard Finnish

>>>>

% supervised systems trained on the normalised corpus outperform those trained on the standardised corpus (WHY?)
% could be because of better word alignment

\begin{table}
  \centering
  \begin{tabular}{|l|r|r|r|}
     \hline
    \textbf{System} & \textbf{PER} & \textbf{WER} & \textbf{BLEU} \\
     \hline
     Colloquial   & 45.31 & 71.28 & 11.95 \\
     Normalised   & 34.63 & 62.34 & 21.81 \\
     Standardised & 38.67 & 63.35 & 17.62 \\
     \hline
     rules-2      & 42.71 & 70.42 & 12.28 \\
     \hline
     cbmt-cn      & 44.44 & 74.17 & 12.35 \\
     pbmt-cn      & \textbf{38.38} & \textbf{66.81} & \textbf{17.90} \\
     \hline
  \end{tabular}
  \caption{Results for the Finnish to English translation task. The first three rows are the results 
    from translating the sections of the corpus.}
  \label{table:results-trad}
\end{table}

The systems for normalisation are able to improve out-of-vocabulary rates in 
many cases, most likely as the online statistical system that we used is trained
on more formal texts. Frequent contractions such as \emph{onks?} `is there?',
\emph{oo} `not' and \emph{vaa} `only' are found untranslated in the output.

@@

We have shown that converting from colloquial Finnish to standard Finnish substantially
helps with the Finnish to English machine translation task.
